---
title: Introduction to ameld
output:
    rmarkdown::html_vignette:
        toc_float: true
vignette: >
    %\VignetteIndexEntry{Introduction to ameld}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
Sys.setenv(LANGUAGE = "en")
library("ameld")
library("survival")

data(eldd)
data(eldr)
```

**Authors**: `r packageDescription("ameld")[["Author"]] `<br />
**Last modified:** `r file.info("ameld.Rmd")$mtime`<br />
**Compiled**: `r date()`

# The `ameld` Package

The `ameld` R package provides a dataset, `eldd`,
of patients evaluated for liver transplantation at the
University Hospital of Leipzig from
`r paste(strftime(attr(eldd, "period"), "%B %Y"), collapse = " to ")`.
`eldr` contains the reference limits for the laboratory measurements used in
`eldd`.

# Access the Data

The datasets could be loaded as follows:

```{r}
library("ameld")
data(eldd)
data(eldr)
```

# Survival Estimates

```{r}
library("survival")

srv <- Surv(eldd$DaysAtRisk, eldd$Deceased)

srvfit <- survfit(srv ~ 1)

plot(
    srvfit,
    mark.time = TRUE,
    main = "Kaplan-Meier Survival Estimate"
)

plot(
    srvfit,
    mark.time = TRUE,
    cumhaz = TRUE,
    main = "Cumulative Hazard"
)

plot(
    srvfit,
    mark.time = TRUE,
    main = "Kaplan-Meier Survival Estimate to day 90",
    xmax = 90
)

plot(
    srvfit,
    mark.time = TRUE,
    cumhaz = TRUE,
    main = "Cumulative Hazard to day 90",
    xmax = 90
)
```

# Acknowledgment

This work is part of the [AMPEL](https://ampel.care/en/)
(Analysis and Reporting System for the Improvement of Patient Safety through
Real-Time Integration of Laboratory Findings) project.

This measure is co-funded with tax revenues based on the budget adopted by
the members of the Saxon State Parliament.

# Session Information

```{r sessionInfo}
sessionInfo()
```
