---
title: Introduction to bootstraping with (a)rcv.gmlnet
output:
    rmarkdown::html_vignette:
        toc_float: true
vignette: >
    %\VignetteIndexEntry{Introduction to (a)rcv.glmnet}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
bibliography:
    - bibliography.bib
    - rpackages.bib
---

```{r setup, include = FALSE}
Sys.setenv(LANGUAGE = "en")
library("ameld")

knitr::write_bib("glmnet", "rpackages.bib")
```

**Authors**: `r packageDescription("ameld")[["Author"]] `<br />
**Last modified:** `r file.info("ameld.Rmd")$mtime`<br />
**Compiled**: `r date()`

# Introduction

The `ameld` R package extends `glmnet::cv.glmnet` [@R-glmnet; @glmnet2010].
It supports a repeated cross-validation (`rcv.glmnet`) and
a repeated cross-validation to tune *alpha* and *lambda* simultaneously
(`arcv.glmnet`). Additionally it provides a `bootstrap` function that could
utilize both functions and supports survival data as described in @harrell1996.

# Dataset

We use the `eldd` dataset provided by `ameld` (see `?eldd` for details) and
standardize it using the `zlog` [@hoffmann2017] method.

```{r dataset}
library("ameld")
library("zlog")
data(eldd)
data(eldr)

# transform reference data.frame for zlog
r <- eldr[c("Code", "AgeDays", "Sex", "LowerLimit", "UpperLimit")]
names(r) <- c("param", "age", "sex", "lower", "upper")
r$age <- r$age / 365.25
r <- set_missing_limits(r)

## we just want to standardize laboratory values
cn <- colnames(eldd)
cnlabs <- cn[grepl("_[SCEFQ1]$", cn)]
zeldd <- eldd
zeldd[c("Age", "Sex", cnlabs)] <- zlog_df(eldd[, c("Age", "Sex", cnlabs)], r)
zeldd[c("Age", "Sex", cnlabs)] <- impute_df(zeldd[c("Age", "Sex", cnlabs)], r)
zeldd <- na.omit(zeldd)
```

# Bootstrapping

Next we apply the bootstrapping. In general the number of bootstrap samples
`nboot` should be equal or larger than 100.
We use a much smaller number here to keep the runtime low.

```{r boot}
library("future")
srv <- Surv(zeldd$DaysAtRisk, zeldd$Deceased)
zeldd$DaysAtRisk <- zeldd$Deceased <- NULL
x <- data.matrix(zeldd)

bt <- bootstrap(
    x, srv,
    fun = rcv.glmnet,
    family = "cox",
    nboot = 3,
    nfolds = 3,
    nrep = 2
)
```

We could show an optimism corrected calibration curve.

```{r plotcal}
plot(bt, what = "calibration")
```

Additionally we could see which variables are selected in each bootstrap step.

```{r plotsel}
plot(bt, what = "selected")
```

## Automatically select best alpha in each Bootstrapping Step.

It is possible to use `arcv.glmnet` to automatically select the best alpha in
each bootstrap step.

```{r eval = FALSE}
selarcv <- function(...) {
    dots <- list(...)
    a <- arcv.glmnet(...)
    i <- which.min.error(a, s = dots$s, maxnnzero = dots$maxnnzero)
    a$models[[i]]
}

bt <- bootstrap(
    x, srv,
    fun = selarcv,
    family = "cox",
    alpha = seq(0, 1, len = 11)^3,
    s = "lambda.1se",
    maxnnzero = 9,
    nboot = 10L, nfolds = 3, nrep = 5,
    m = 50, times = 90
)
```

# Acknowledgment

This work is part of the [AMPEL](https://ampel.care/en/)
(Analysis and Reporting System for the Improvement of Patient Safety through
Real-Time Integration of Laboratory Findings) project.

This measure is co-funded with tax revenues based on the budget adopted by
the members of the Saxon State Parliament.

# Session Information

```{r sessionInfo}
sessionInfo()
```

# References
